Low Complexity Adaptive Sonar Imaging

### 2016-03-02 LCAforActiveSonarImaging/lca.tex
la2pdf commit - 2016:03:02	13:35:59	(on. 2.mars)

To build confidence in the LCA method we show that a robust MVDR implementation typically creates weights spanning between the rectangular and Hamming window function. We let LCA select from a set of Kaiser windows with responses in this span, and add some steered variations of each. We limit the steering to roughly half the -3\,dB width of the window's amplitude response. Using experimental data from the Kongsberg Maritime HISAS 1030 sonar we find that LCA and MVDR produce nearly identical images of large scenes, both being superior to DAS. On point targets LCA is able to double the resolution compared to DAS, or provide half that of MVDR. This performance is achieved with a total of 6 windows; the rectangular and Kaiser window with $\beta=5$, in unsteered version, and left and right steered to the steering limit. Slightly smoother images are produced if the window count is increased to 15, but past this we observe minimal difference. Finally we show that LCA works just as well if Kaiser windows are substituted with trigonometric ones.

@@

A much faster and simpler alternative to MVDR is the Low Complexity Adaptive (LCA) beamformer. Based on an idea by Vignon~\cite{Vignon2008}, it was first introduced by Synnev\aa{}g \emph{et al}. in clinical medical imaging~\cite{Synnevag2008}, who demonstrated its ability to obtain very similar image quality to MVDR in systems with focused transmit beams. LCA applies a set of predefined windows, or apodization weights, and selects the one offering the best sidelobe suppression. It relies on the same optimization criterion as MVDR, thus may be considered a version of MVDR with a discrete and static window solution space. Alternatively, it may be viewed as a multi-apodization technique. One such method is described by Stankwitz \emph{et al}. in radar imaging~\cite{Stankwitz1995}, where the best out of 2 or 3 windows are selected. However, LCA is different in that it typically selects from a larger pool of windows and allows micro-steered variations of them. This gives LCA an adaptive freedom that is somewhere in between that of traditional multi-apodization techniques and MVDR. The bigger the space the weights span, the greater the adaptive freedom, but only the solutions that are valid in the constrained solution space can make a difference.

Synnev\aa{}g \emph{et al}. suggested using a window set comprised of rectangular, Kaiser (or Kaiser-Harris) and inverted Kaiser functions. In total he made LCA choose from 12 unique windows, 6 of which were phase shifted versions of a fairly wide Kaiser window, and one rectangular. Phase shifting the windows has the effect of slighly adjusting the angle the array is steered towards. His composition of windows was motivated by the desire to use wide windows with maximum sidelobe suppression near the receiver, and narrow windows at greater depths for maximum sensitivity and depth penetration. While achieving good results, it was not made clear how many windows LCA needs, or how they ideally should be steered. Also, it is not apparent how LCA performs in sonar systems where the transmit beam is unfocused.

In this study we expand our previous work of applying LCA to active sonar imaging~\cite{Buskenes2011}. We use variations of the parametric Kaiser window, but also compare it to the trigonometric window function. We extend Synnev\aa{}g's work by steering all our windows and not just some, and demonstrate how LCA's performance is affected by the type and number of windows included. Our results are obtained using experimental data from the 32 element Kongsberg Maritime HISAS 1030 synthetic aperture sonar (SAS), which operates at 100\,kHz with an -3\,dB element opening angle of 23$^\circ$.

@@

Since pixels are correlated laterally, we can include a weighted combination of these as well to improve the variance estimation:
%
\begin{align}
s_z^2 = \frac{1}{N_x N_k} \sumb{\nmat{\scriptstyle x'=\\[-2mm]\scriptstyle x-X}}{x+X} \sumb{\nmat{\scriptstyle n'=\\[-2mm]\scriptstyle n-K}}{n+K} \omega[x',n']\big| z[x',n'] \big|^2
\end{align}
%
where $N_x = 2X+1$ is the number of azimuth lines to average over, and $\omega[x',n']$ is a normalized 2 dimensional weight function. However, we observed no visual improvement doing this, and the effect of micro-steered windows is hard to predict. Henceforth no lateral averaging will be applied.
% 
% \begin{itemize}
% \item \emph{Window function}
% \item \emph{Parameter boundaries}
% \item \emph{Number of windows needed}
% \item \emph{Oversampling need}
% \item \emph{Image quality}
% \end{itemize}

@@

To test the performance of the MVDR and LCA beamformers, we have processed data acquired by the Kongsberg Maritime HUGIN AUV carrying their 32 element HISAS1030 sonar~\cite{Hansen2011}. It is a high resolution synthetic aperture sonar with 1.2\;m array length, 100\;kHz operating frequency, 30\;kHz bandwidth and 23$^\circ$ element -3\,dB opening angle.

>>>>

To build confidence in the LCA method we show that a robust MVDR implementation typically creates weights spanning between the rectangular and Hamming window function. We let LCA select from a set of Kaiser windows with responses in this span, and add some steered variations of each. We limit the steering to roughly half the -3\,dB width of the window's amplitude response. Using experimental data from the Kongsberg Maritime HISAS1030 sonar we find that LCA and MVDR produce nearly identical images of large scenes, both being superior to DAS. On point targets LCA is able to double the resolution compared to DAS, or provide half that of MVDR. This performance is achieved with a total of 6 windows; the rectangular and Kaiser window with $\beta=5$, in unsteered version, and left and right steered to the steering limit. Slightly smoother images are produced if the window count is increased to 15, but past this we observe minimal difference. Finally we show that LCA works just as well if Kaiser windows are substituted with trigonometric ones.

@@

A much faster and simpler alternative to MVDR is the Low Complexity Adaptive (LCA) beamformer. Based on an idea by Vignon~\cite{Vignon2008}, it was first introduced by Synnev\aa{}g \emph{et al}. in clinical medical ultrasound imaging~\cite{Synnevag2008}, who demonstrated its ability to obtain very similar image quality to MVDR in systems with focused transmit beams. LCA applies a set of predefined and static windows, or apodization weights, and selects the one offering the best noise suppression. It relies on the same optimization criterion as MVDR, thus may be considered a version of MVDR with a discrete and static window solution space. Alternatively, it may be viewed as a multi-apodization technique. One such method is described by Stankwitz \emph{et al}. in radar imaging~\cite{Stankwitz1995}, where the best out of 2 or 3 windows are selected. However, LCA is different in that it typically selects from a larger pool of windows and allows phase-steered variations of them. This gives LCA an adaptive freedom that lies somewhere in between that of traditional multi-apodization techniques and MVDR. How similarly LCA performs compared to MVDR depends on how well the predefined windows represent the constrained solution space of MVDR.

% While using weights that span a bigger space gives LCA greater freedom, not The bigger the space the weights span, the greater the adaptive freedom, but only the solutions that are valid in the constrained solution space can make a difference.

% Instead of adaptively computing the weights, it adaptively selects them.

Synnev\aa{}g \emph{et al}. suggested using a window set comprised of rectangular, Kaiser (or Kaiser-Harris) and inverted Kaiser functions. In total he made LCA choose from 12 unique windows, 6 of which were phase-shifted versions of a fairly wide Kaiser window, and one were rectangular. Phase-shifting the windows has the effect of slighly adjusting the angle the array is steered towards. Synnev\aa{}g's composition of windows was motivated by the desire to use wide windows with maximum sidelobe suppression near the receiver, and narrow windows at greater depths for maximum sensitivity and depth penetration. While achieving good results, it was not made clear how many windows LCA needs, or how they ideally should be steered. Also, it is not apparent how LCA performs in sonar systems where the transmit beam is unfocused.

In this study we expand our previous work of applying LCA to active sonar imaging~\cite{Buskenes2011}. We use variations of the parametric Kaiser window, but also compare it to the trigonometric window function. We extend Synnev\aa{}g's work by steering all our windows and not just some, and demonstrate how LCA's performance is affected by the type and number of windows included. Our results are obtained using experimental data from the 32 element Kongsberg Maritime HISAS1030, which operates at 100\,kHz with an -3\,dB element opening angle of 23$^\circ$. Although this system was designed as synthetic aperture sonar (SAS) where an image is synthesized from several pings of data, we will only create images out of single pings (sectorscan imaging) in this study.

@@

Since pixels are correlated laterally, we can include a weighted combination of these as well to improve the variance estimation:
%
\begin{align}
s_z^2 = \frac{1}{N_x N_k} \sumb{\nmat{\scriptstyle x'=\\[-2mm]\scriptstyle x-X}}{x+X} \sumb{\nmat{\scriptstyle n'=\\[-2mm]\scriptstyle n-K}}{n+K} \omega[x',n']\big| z[x',n'] \big|^2
\end{align}
%
where $N_x = 2X+1$ is the number of azimuth lines to average over, and $\omega[x',n']$ is a normalized 2 dimensional weight function. However, we observed no visual improvement doing this, and the effect of phase-steered windows is hard to predict. Henceforth no lateral averaging will be applied.
% 
% \begin{itemize}
% \item \emph{Window function}
% \item \emph{Parameter boundaries}
% \item \emph{Number of windows needed}
% \item \emph{Oversampling need}
% \item \emph{Image quality}
% \end{itemize}

@@

To test the performance of the MVDR and LCA beamformers, we have processed data acquired by the Kongsberg Maritime HUGIN AUV carrying their 32 element HISAS1030 sonar~\cite{Hansen2011}. It is a high resolution sonar with 1.2\;m array length, 100\;kHz operating frequency, 30\;kHz bandwidth and 23$^\circ$ element -3\,dB opening angle.

### 2016-03-02 LCAforActiveSonarImaging/submission/submit1/answer/answer.tex
la2pdf commit - 2016:03:02	13:35:59	(on. 2.mars)

\begin{enumerate}
% \emph{General}
\hi{Abstract}
\qi{"the array's spatial covariance matrix" might be a bit shocking for the non-expert Reader $\rightarrow$ The formulation in the Introduction might fit better "the influence of noise ..."}
\ai{This is a good point. We have reformulated the sentence to speak of computing the spatial statistics instead.}
%
\hi{Page 2}
\qi{"6 of which were micro-steered" $\rightarrow$ a short explanation or reference might help to understand what is meant by "micro-steered"}
\ai{Another good point. This wording is now changed and a sentence elaborating on its meaning is added.}
%
\qi{"it was not clear how how many" $\rightarrow$ delete 1 x "how"}
\ai{Thank you. Corrected.}
%
\qi{"with an element beamwidth" $\rightarrow$ what is meant by element beamwidth? Maybe an additional figure could be used to explain this.}
\ai{Good remark. We meant the -3dB opening angle of the elements. We have reworded this the two places it is mentioned.}
%
\hi{Page 3, after (1)}
\qi{"Applying a window that trails off .... Minimum variance distortion-less response (MVDR) beamformer." $\rightarrow$ real and complex Windows? The introduction should be more formal: x ist introduced as delayed data, weights or delays to change, or both is explained later. $\rightarrow$ The difference between noise and sidelobes should be pointed out.}
\ai{}
%
\hi{Page 5, equation (5)}
\qi{The lateral correlation and the reason why equation (5) is a good choice should be better explained.}
\ai{Indeed it should, this paragraph is now rewritten.}
%
\hi{Page 6, last row}
\qi{"they are optimal for any value" $\rightarrow$ optimal in which sense?}
\ai{Thank you. This was mentioned earlier in the same paragraph, just after (8).}
%
\hi{Fig 3}
\qi{why is a 40x40 pixel group a good choice? could the associated broadside direction be made visible in the figure, too?}
\ai{This is an excellent question. This should perhaps be related to the detail level in the image, which in turn depends on the array's imaging performance.}
%
\hi{Page 8}
\qi{reference to Fig.1 $\rightarrow$ could the sequence of figures be changes to have Fig. 1 closer to this paragraph}
\ai{Valid point. In our two column print the figure is on the page before its first reference. }
%
\hi{Page 11}
\qi{where is the synthetic aperture needed/applied in this work, or is it just the name of the array? how does the work on LCA/MVDR Change the Quality of synthetic aperture?
what exactly is the sectorscan mode?}
\ai{Thank you, we see that our formulation should be clarified. We are creating an image of the sector highlighted by a single ping of data (sectorscan), but are using data recorded by a synthetic aperture sonar (SAS). We mention it since SAS systems have a wide transmit beam intended to highlight multiple pings and thus does not produce the best sectorscan images.}
%
\end{enumerate}

>>>>

\begin{enumerate}
% \emph{General}
\hi{Abstract}
\qi{"the array's spatial covariance matrix" might be a bit shocking for the non-expert Reader $\rightarrow$ The formulation in the Introduction might fit better "the influence of noise ..."}
\ai{This is a good point. We have reformulated the sentence to speak of computing the spatial statistics instead.}
%
\hi{Page 2}
\qi{"6 of which were micro-steered" $\rightarrow$ a short explanation or reference might help to understand what is meant by "micro-steered"}
\ai{Another good point. This wording is now changed and a sentence elaborating on its meaning is added.}
%
\qi{"it was not clear how how many" $\rightarrow$ delete 1 x "how"}
\ai{Thank you. Corrected.}
%
\qi{"with an element beamwidth" $\rightarrow$ what is meant by element beamwidth? Maybe an additional figure could be used to explain this.}
\ai{Good remark. We meant the -3dB opening angle of the elements. We have reworded this the two places it is mentioned.}
%
\hi{Page 3, after (1)}
\qi{"Applying a window that trails off .... Minimum variance distortion-less response (MVDR) beamformer." $\rightarrow$ real and complex Windows? The introduction should be more formal: x ist introduced as delayed data, weights or delays to change, or both is explained later. $\rightarrow$ The difference between noise and sidelobes should be pointed out.}
\ai{}
%
\hi{Page 5, equation (5)}
\qi{The lateral correlation and the reason why equation (5) is a good choice should be better explained.}
\ai{Indeed it should, this paragraph is now rewritten.}
%
\hi{Page 6, last row}
\qi{"they are optimal for any value" $\rightarrow$ optimal in which sense?}
\ai{Thank you. This was mentioned earlier in the same paragraph, just after (8).}
%
\hi{Fig 3}
\qi{why is a 40x40 pixel group a good choice? could the associated broadside direction be made visible in the figure, too?}
\ai{This is an excellent question. This should perhaps be related to the detail level in the image, which in turn depends on the array's imaging performance.}
%
\hi{Page 8}
\qi{reference to Fig.1 $\rightarrow$ could the sequence of figures be changes to have Fig. 1 closer to this paragraph}
\ai{Valid point. In our two column print the figure is on the page before its first reference. }
%
\hi{Page 11}
\qi{where is the synthetic aperture needed/applied in this work, or is it just the name of the array? how does the work on LCA/MVDR Change the Quality of synthetic aperture? What exactly is the sectorscan mode?}
\ai{The reviewer is quite right, it is a system designed for SAS imaging, but we only create image sectors for its data. That is, the images are from single ping, rather than synthesized from several pings. We agree this was worded in an ambiguous way and have now made relevant changes, as well as added a remark of what it meant by sectorscan.}
%
\end{enumerate}
